Lattice Boltzmann simulations of magnetically responsive\\ ellipsoidal particle stabilized emulsions

### 2025-03-20 chapters/paper1.tex
Changes 1. Updated aim 1 with first person wording 2. Updated aim 2 with first person wording and modified conclusion 3. Updated aim 3 with full size conclusion

In this chapter, we investigate the effect of external magnetic fields on the formation of bijels stabilized by 
anisotropic magnetic particles. The microstructure of bijels stabilized with oblate, spherical and prolate particles, 
namely the characteristic length scale and tortuosity, will be assessed as a function of the applied magnetic field
strength. We also characterize the underlying mechanisms governing the observed behavior.

@@

Bijels can be used as emulsion templates for porous materials with
applications including drug delivery, water desalination, and battery
electrodes
\cite{vanoli_bijels_2022, chen_pore-scale_2022, lu_controllable_2020, garcia_scalable_2019}.
The transport properties of porous materials depend on the porosity and
tortuosity of the void space and can be linked to the phase-separated
morphology of the bijel template. Therefore, we study the domain size
and tortuosity of the phase morphology of bijels stabilized by
anisotropic particles. In particular, we investigate the effect of
magnetic particles on the microstructure of bijels that are formed in
external magnetic fields. We simulated bijels with a particle volume
fraction of \(\phi_p=0.15\) and varied the magnetic flux density such
that the Bond numbers varies in the range \(0\le\bar{B}\le1\).
In this work, we use a fixed particle volume fraction and focus on the dependence of the 
bijel structure on the magnetic field. Bijels with varying particle loading and surface tension 
and have been studied elsewhere in the literature, e.g., by Hijnen et al.\cite{hijnen_bijels_2015} 
and Jansen and Harting\cite{jansen_bijels_2011}.

@@

In LB simulations of spinodal decomposition, a characteristic length
scale can be obtained from the structure factor, defined in Section \ref{section:domain_size}. \cite{kendon_3d_1999,kendon_inertial_2001} 
It is worth noting that this definition of average domain size does not necessarily yield values that coincide with the typical
domain size observed visually in snapshots such as those shown in Fig.~\ref{fig:microstructure_viz}. However, in a dynamic
scaling regime, different measures of domain size are expected to differ only by a constant ratio. We found that the commonly 
used measure \(L_1(t)\) tends to overestimate the observed domain size. Numerically, the calculation of the spherically 
averaged structure factor is subject to poor statistics in the low \(k\)-shells, where the average \(|\vec{k}|\) of the 
lattice sites differs from the nominal shell radius. Additionally, the measurement of \(L(t)\) in simulations with
periodic boundary conditions is subject to finite size effects. More importantly, in suspensions of anisotropic particles, 
we cannot expect that the structure factor is isotropic. Therefore, we also used the 3D structure factor to calculate a lateral domain 
size in each Cartesian direction from the second moments also defined in Section \ref{section:domain_size}. 
\cite{jansen_bijels_2011,gunther_timescales_2014} We add the additional processing calculating a parallel and perpendicular domain size
from the results in each cartesian direction, defining $L_{\parallel}=L_z$ and $L_{\perp} = \frac{L_x+L_y}{2}$

Before calculating the structure factor \(S(\vec{k},t)\), we first filled the lattice sites covered by particles with the average
density at the surrounding lattice sites as detailed in Section \ref{section:filling_routine}. We plot the structure factor of the structures in
Figure \ref{fig:structure_factor}.

@@

At first glance, the time-dependence of the domain size in Fig.
\ref{fig:domain_size} suggests that the domain size is not strongly
affected by the external magnetic field. In Fig.\ref{fig:D2a_B}a) we
plotted the domain size measure \(L_d\) normalized by \(R\), the larger
of the two radii of the ellipsoidal particles. The data points represent
the domain size measured at the final time step and averaged over three
independent simulation runs with the same parameters. The normalization
with the particle radius makes it apparent that the ellipsoidal
particles reduce the domain size of bijels. This observation is in line
with observations by Günther et al.~\cite{gunther_timescales_2014} and
indicates that -- relative to their volume -- ellipsoidal particles
stabilize a larger interface area than spherical particles. This is due
to the larger cross-sectional area of ellipsoidal particles at the same
volume. Furthermore, steric constraints prevent anisotropic particles to
pack less densely in the interface than spherical particles, as we will
discuss in more detail below.

The magnetic field does have an effect on the directional
domain size as shown in Figure \ref{fig:D2a_B} where we plotted the
domain size in the direction parallel and perpendicular to \(\vec{B}\)
separately. We find that for prolate particles, the domain size in the
parallel direction increases with the field strength up to approximately
three particle radii for the largest field, while the perpendicular
domain size stays constant. This trend is reversed for oblate particles
for which the domain size in the parallel direction decreases slightly
with field strength while the perpendicular domain size increases by
approximately four particle radii. Without a field, prolate
particles lead to smaller domain size than both spherical and oblate
particles, whereas they increase the domain size when the bijel forms under the
influence of a magnetic field. The anisotropy is due to the alignment of
ellipsoidal particles which leads to a closer packing within the
interface akin to nematic ordering. The difference between oblate and
prolate particles is caused by the orientation of the magnetic dipole
moment \(\vec{m}\) along the symmetry axis of the particles. In our
model, \(\vec{m}\) is aligned with the larger axis of prolate
ellipsoids, whereas it is aligned with the smaller axis of oblate
ellipsoids. Hence \(\vec{m}\) lies in the plane of the larger
cross-section of prolate particles while \(\vec{m}\) is perpendicular to
the larger cross-section of oblate particles. We thus expect that
alignment of the particles with the magnetic field \(\vec{B}\)
facilitates closer packing perpendicular to the field for prolate
particles, and parallel to the field for oblate particles. The
increase/decrease of domain size levels off at intermediate field
strength and further increases of the field have no additional effect.

@@

The anisotropy of domain size suggests that the magnetic field
influences the bijel morphology at the microscale. To investigate
whether these structural changes have an effect on macroscopic
transport properties, we also characterized the tortuosity of the
bijels. Various definitions of tortuosity have been proposed in the
literature \cite{dasilva_tortuosity_2022}, including geometric,
diffusional, and hydraulic tortuosity. Here we use the diffusional
tortuosity $\tau= \frac{\epsilon D}{D_{\text{eff}}}$, i.e., the ratio of the 
effective diffusivity $D_{\text{eff}}$ and the intrinsic diffusivity
$D$. We calculate the tortuosity using the procedure detailed in Section \ref{section:tortuosity}

Figure \ref{fig:tau_B} shows the results for the tortuosity as a
function of the applied magnetic flux for the three particle aspect
ratios. In the absence a magnetic field, the three aspect ratios lead to
a similar tortuosity around $\tau\approx 1.5$, slightly larger than
the tortuosity predicted by the Bruggeman relation
\(\tau=\epsilon^{-0.5}=1.41\) for equal volume fractions
\(\epsilon=0.5\) of the phases
\cite{bruggeman1935tortuosity, tjaden_origin_2016}. The observed
tortuosity is consistent with simulations of gyroid structures by Luo et
al., who found diffusive tortuosities in the range from 1.48 to 1.73
\cite{luo_macroscopic_2020}. Figure \ref{fig:tau_B} shows that in an
applied magnetic field, the tortuosity of bijels stabilized by
anisotropic particles (\(\alpha\neq1\)) becomes anisotropic as well.
Oblate particles (\(\alpha=0.5\)) lead to an increase of the tortuosity
\(\tau_\parallel\) in the direction of the magnetic field while the
tortuosity in the direction perpendicular to the magnetic field remains
around \(\tau_\parallel\approx1.5\). Conversely, prolate particles
(\(\alpha=2\)) lead to an increase of the tortuosity in the direction
perpendicular to the magnetic field, while the tortuosity in the
parallel direction slightly decreases with increasing field strength. In
both cases, the changes in tortuosity are consistent with the
anisotropic domain size and the alignment of particles with the magnetic
field. For prolate particles, the longer axis is aligned with the
magnetic moment and induces larger domain size and lower tortuosity in
this direction. The results show that the magnetic field has a
noticeable effect on the tortuosity of the bijel. We observe that the
largest change of tortuosity is measured at intermediate field strength
\(\bar{B}=0.5\). This suggests that the competition of interfacial and
magnetic forces and the resulting alignment of particles and liquid
domains induces morphological changes that are more complex than a
monotonic increase or decrease. We therefore turn to investigate the
microstructural changes within the bijel morphology in more detail.

@@

Beyond the length scale, curvature analysis offers insights into the
local topological landscape of the bijel, facilitating further
microstructure optimization for various applications.
\cite{reeves_quantitative_2016} We plot the area averaged mean and
gaussian curvature \(H\Sigma^{-1}\) and \(K\Sigma^{-2}\) for bijels
stabilized with oblate, spherical and prolate ellipsoids in Figure
\ref{fig:curvature-vs-B_ss}.

\begin{figure} 
    \centering 
    \includegraphics[scale = 0.4]{figures/results/paper1/curvature-vs-B_ss.png} 
    \caption{Plot of the area averaged mean $\langle H \rangle \Sigma^{-1}$ and Gaussian 
            $\langle K \rangle \Sigma^{-2}$ curvature in the left and right respectively. Each 
            plot contains the $\langle H \rangle \Sigma^{-1}$ and $\langle K \rangle \Sigma^{-1}$ 
            at the final timestep averaged across 3 runs for particles with $\alpha = 0.5, 1, 2$. 
            We observe that the mean curvature does not demonstrate any trends, while the Gaussian 
            curvature shows a reduction as the magnetic field strength is increased for 
            ellipsoidal particles.} 
    \label{fig:curvature-vs-B_ss}
\end{figure}

Error bars are represented by the standard
deviation of the three simulations per parameter set performed. The area
averaged mean curvature shows little dependence on the applied field
strength with values around 0 and within error of one another. A value
near or at 0 is expected as the fluid volume fraction is equal and the
particles are neutrally wetting, implying the interface should not have
a preferred direction of curvature. \cite{jinnai_interfacial_2001} For
ellipsoidal particles, the gaussian curvature becomes more negative as
larger field strengths are applied. We also see that the gaussian
curvature becomes more positive as the interfacial area of the particle
increases. The trend of these results are inverse to the interfacial
area of each particle, \(A_{rm}\). This is expected behavior and is in
agreement with experiments that have investigated the effect of particle
size on the curvature of bijels. \cite{reeves_quantitative_2016}

Reeves and coworkers identified that nanoparticles were better able to
lock in the spinodal structure of the liquid-liquid phase separation as
they were less disruptive the the curvature of the interface due to
their lower interfacial adsorption energy. This effect can be observed
as a more negative Gaussian curvature as a result of the greater number
of hyperbolic surfaces, a hallmark of a minimal surface. In our system,
the oblate and prolate particles have 60\% and 30\% more interfacial
area than the spherical particles, resulting in the ellipsoidal
particles being more disruptive to the curvature of the interface,
observed as a more positive Gaussian curvature. As the magnetic field
strength is increased, the packing of the particles at the interface
changes, as observed from the radial distribution function.
\cite{karthikeyan_formation_2024} These changes in particle packing
change the way the particles affect the curvature of the interface,
resulting in the curvature changes observed. To investigate the field
dependence on the curvature, we plot the area averaged Gaussian
curvature as a function of time.

\begin{figure} 
    \centering 
    \includegraphics[scale = 0.4]{figures/results/paper1/genus_curvature_vs_coverage.png} 
    \caption{Plots of Gaussian $\langle K \rangle \Sigma^{-2}$ curvature over time averaged across 3 runs. 
    From left to right, we have particles with $\alpha = 0.5, 1, 2$. We observe that the initial stages of 
    evolution are unaffected by the applied magnetic field. However, after about 10000 timsteps, we see a 
    magnetic field dependence on the evolution of the Gaussian curvature. The preservation of this divergence 
    occurs when ellipsoidal particles are used but not spherical particles.} 
    \label{fig:curvature_vs_coverage}
\end{figure}

Given we did not see any noteworthy changes in the mean curvature in
Figure \ref{fig:curvature-vs-B_ss}, we focused on the Gaussian curvature
in Figure \ref{fig:curvature_vs_coverage}. We observe that for all
simulated systems, the Gaussian curvature increases over time until an
invariance is reached near the end of the simulation, signifying the
bijel microstructure is locked in place. We also observe that for all
microstructures the curvature over time does not increase at the same
rate, characterized with dips at 10 to 20,000 timesteps for all
particles and a magnetic field dependent change in the curvature
increase.

However, for the spherical particles this local topology change
disappears once the spinodal decomposition ceases, observed as a
convergence of all curvatures. This is not observed for the bijels
stabilized by ellipsoidal particles. Despite all our curvature
calculations being normalized with their surface area to volume ratio,
we still see curvature differences as a function of the applied magnetic
field. Experiments have identified that curvature changes can occur due
to the formation of monogels, which are particle networks formed from
bijels that can withstand the remixing of their constituent liquids.
\cite{sanz_colloidal_2009, lee_making_2013} This process cannot occur in
our simulations as we are utilizing repulsive particles. Therefore, the
change in curvature arises from the differences in local state of the
colloids at the interface. From the microstructure results in Figures \ref{fig:D2a_B}
in addition to the genus differences characterized in Figure \ref{fig:curvature_vs_coverage},
the connectivity of the microstructure is affected through the application of magnetic fields.
We can use the channel size distribution to investigate how the connectivity of the microstructure and thus the 
channel size is affected as a function of the applied magnetic field.

While minor run to run variation is observed, the combined data of the channel size distribution shown in Figure \ref{fig:CSD_B_ss} shows some major differences 
in the channel size distribution in each bijel stabilized with oblate particles. It was expected that the domain anisotropy obtained using second moment of the 
structure factor would manifest as two peaks in the channel size distribution. \cite{karthikeyan_formation_2024} This behavior is observed in the channel size 
distribution for bijels stabilized with oblate and prolate particles with $\bar{B} = 1$, but is more difficult to see when smaller magnetic fields are applied. 
We see that there is a shift to larger channel sizes upon application of the magnetic field for bijels stabilized by ellipsoids. These differences arise from the 
particle ordering and local distribution changes observed.

\begin{figure} 
        \centering 
        \includegraphics[scale = 0.4]{figures/results/paper1/CSD_ss.png} 
        \caption{Plots of the channel size distribution for bijels stabilized by oblate, spherical and oblate particles at the final timestep plotted
                 against the distance from the interface normalized with the characteristic length scale. The channel sizes obtained are averaged over 
                 three runs. We see that the peak locations are shifted to larger and smaller positions for bijels stabilized with ellipsoidal particles
                 upon field application} 
        \label{fig:CSD_B_ss}
    \end{figure}

@@

To shed light on the mechanisms involved in the coarsening dynamics,
we now turn to the kinetics of bijel formation in more detail. The
coarsening dynamics can be characterized by a coarsening speed. Here,
we use the finite difference of the directional domain size to
calculate the components of the coarsening velocity
%
\begin{equation}
u_{L_\beta}(t) = \frac{L_\beta(t+\Delta t_s)-L_\beta(t-\Delta t_s)}{2\Delta t_s} ,
\end{equation}
%
where \(\Delta t_s\) is the time between simulation
snapshots, and $\beta$ denotes either the direction parallel or perpendicular to the magnetic field.

@@

Figure \ref{fig:coarsening_velocity} shows the coarsening speed in the directions parallel and perpendicular to the applied magnetic field. The data confirms that for spherical particles, domain coarsening remains isotropic. We observe a higher coarsening speed initially which slows at later times when the particles begin to jam. This indicates that the
coarsening is subject to different time scales, as reported previously
by Harting and co-workers \cite{gunther_timescales_2014}. Reeves et
al.~\cite{reeves_particle-size_2015} have pointed out that bijel
formation hinges on the jamming time in relation to the disruption time, i.e., the time scale at which domain pinch-off events can cause bijel break-up through depercolation. In our simulations, however, we have not observed disruption of bijel formation; stable bijels form for all parameters consistent with simulations by Stratford et
al.~\cite{stratford_colloidal_2005} and Jansen et al.~\cite{jansen_bijels_2011}.

For ellipsoidal particles, our data clearly shows that an\-isotropic
domain coarsening arises in magnetic fields.  For oblate particles
(\(\alpha=0.5\)), the coarsening speed in the direction perpendicular
to the field increases with increasing magnetic field strength.
Jamming in this direction appears to be delayed compared with the
direction parallel to the field. The coarsening speed in the parallel
direction remains comparable to the case without an applied magnetic
field. This anisotropic behavior of the coarsening speed reverses for
prolate particles, where the coarsening speed in the direction parallel to the field increases with increasing field strength, and jamming in this direction
is delayed compared with the perpendicular direction. In addition, we observe a shoulder (around \(10^4\Delta t\)) where the coarsening speed decays more slowly before jamming sets in. These results suggest that the coarsening near the jamming time is affected by a mechanism that depends on the applied magnetic field and causes the anisotropic
behavior. This mechanisms involves the re-orientation of anisotropic
particles and their alignment relative to the direction of the
magnetic field. To confirm this hypothesis, we analyze the
orientational order of the particles and their alignment relative to
the magnetic field and the interface, respectively.

@@

We select $\langle Q6 \rangle$ and $\langle Q2 \rangle$ as the bond order parameters to calculate as we are interested in identifying the point when ordering 
of the material changes. Kapfer et al. demonstrated that only using $\langle Q6 \rangle$ to identify the ordering of particles in the system can result in 
false positives when the particles have icosahedral or amorphous structure. \cite{kapfer_jammed_2012} Mickel et al. showed how increasing deviations from 
$\langle Q2 \rangle = 0$ can be used to characterize increasing disorder in a system. \cite{mickel_shortcomings_2013} By plotting $\langle Q6 \rangle$ and 
$\langle Q2 \rangle$, we will be able to identify changes in the local ordering of the system even if we cannot directly assess the local structure.

>>>>

In this chapter, I investigate the effect of external magnetic fields on the formation of bijels stabilized by 
anisotropic magnetic particles. The microstructure of bijels stabilized with oblate, spherical and prolate particles, 
namely the characteristic length scale and tortuosity, will be assessed as a function of the applied magnetic field
strength. I also characterize the underlying mechanisms governing the observed behavior.

@@

Bijels can be used as emulsion templates for porous materials with
applications including drug delivery, water desalination, and battery
electrodes
\cite{vanoli_bijels_2022, chen_pore-scale_2022, lu_controllable_2020, garcia_scalable_2019}.
The transport properties of porous materials depend on the porosity and
tortuosity of the void space and can be linked to the phase-separated
morphology of the bijel template. Therefore, I study the domain size
and tortuosity of the phase morphology of bijels stabilized by
anisotropic particles. In particular, I investigate the effect of
magnetic particles on the microstructure of bijels that are formed in
external magnetic fields. I simulated bijels with a particle volume
fraction of \(\phi_p=0.15\) and varied the magnetic flux density such
that the Bond numbers varies in the range \(0\le\bar{B}\le1\).
In this work, I use a fixed particle volume fraction and focus on the dependence of the 
bijel structure on the magnetic field. Bijels with varying particle loading and surface tension 
and have been studied elsewhere in the literature, e.g., by Hijnen et al.\cite{hijnen_bijels_2015} 
and Jansen and Harting\cite{jansen_bijels_2011}.

@@

In LB simulations of spinodal decomposition, a characteristic length
scale can be obtained from the structure factor, defined in Section \ref{section:domain_size}. \cite{kendon_3d_1999,kendon_inertial_2001} 
It is worth noting that this definition of average domain size does not necessarily yield values that coincide with the typical
domain size observed visually in snapshots such as those shown in Fig.~\ref{fig:microstructure_viz}. However, in a dynamic
scaling regime, different measures of domain size are expected to differ only by a constant ratio. I found that the commonly 
used measure \(L_1(t)\) tends to overestimate the observed domain size. Numerically, the calculation of the spherically 
averaged structure factor is subject to poor statistics in the low \(k\)-shells, where the average \(|\vec{k}|\) of the 
lattice sites differs from the nominal shell radius. Additionally, the measurement of \(L(t)\) in simulations with
periodic boundary conditions is subject to finite size effects. More importantly, in suspensions of anisotropic particles, 
I cannot expect that the structure factor is isotropic. Therefore, I also used the 3D structure factor to calculate a lateral domain 
size in each Cartesian direction from the second moments also defined in Section \ref{section:domain_size}. 
\cite{jansen_bijels_2011,gunther_timescales_2014} I add the additional processing calculating a parallel and perpendicular domain size
from the results in each cartesian direction, defining $L_{\parallel}=L_z$ and $L_{\perp} = \frac{L_x+L_y}{2}$ I plot the structure factor of the structures in
Figure \ref{fig:structure_factor}.

@@

At first glance, the time-dependence of the domain size in Fig.
\ref{fig:domain_size} suggests that the domain size is not strongly
affected by the external magnetic field. In Fig.\ref{fig:D2a_B}a) I
plotted the domain size measure \(L_d\) normalized by \(R\), the larger
of the two radii of the ellipsoidal particles. The data points represent
the domain size measured at the final time step and averaged over three
independent simulation runs with the same parameters. The normalization
with the particle radius makes it apparent that the ellipsoidal
particles reduce the domain size of bijels. This observation is in line
with observations by Günther et al.~\cite{gunther_timescales_2014} and
indicates that -- relative to their volume -- ellipsoidal particles
stabilize a larger interface area than spherical particles. This is due
to the larger cross-sectional area of ellipsoidal particles at the same
volume. Furthermore, steric constraints prevent anisotropic particles to
pack less densely in the interface than spherical particles, as I will
discuss in more detail below.

The magnetic field does have an effect on the directional
domain size as shown in Figure \ref{fig:D2a_B} where I plotted the
domain size in the direction parallel and perpendicular to \(\vec{B}\)
separately. I find that for prolate particles, the domain size in the
parallel direction increases with the field strength up to approximately
three particle radii for the largest field, while the perpendicular
domain size stays constant. This trend is reversed for oblate particles
for which the domain size in the parallel direction decreases slightly
with field strength while the perpendicular domain size increases by
approximately four particle radii. Without a field, prolate
particles lead to smaller domain size than both spherical and oblate
particles, whereas they increase the domain size when the bijel forms under the
influence of a magnetic field. The anisotropy is due to the alignment of
ellipsoidal particles which leads to a closer packing within the
interface akin to nematic ordering. The difference between oblate and
prolate particles is caused by the orientation of the magnetic dipole
moment \(\vec{m}\) along the symmetry axis of the particles. In our
model, \(\vec{m}\) is aligned with the larger axis of prolate
ellipsoids, whereas it is aligned with the smaller axis of oblate
ellipsoids. Hence \(\vec{m}\) lies in the plane of the larger
cross-section of prolate particles while \(\vec{m}\) is perpendicular to
the larger cross-section of oblate particles. I thus expect that
alignment of the particles with the magnetic field \(\vec{B}\)
facilitates closer packing perpendicular to the field for prolate
particles, and parallel to the field for oblate particles. The
increase/decrease of domain size levels off at intermediate field
strength and further increases of the field have no additional effect.

@@

The anisotropy of domain size suggests that the magnetic field
influences the bijel morphology at the microscale. To investigate
whether these structural changes have an effect on macroscopic
transport properties, I also characterized the tortuosity of the
bijels. Various definitions of tortuosity have been proposed in the
literature \cite{dasilva_tortuosity_2022}, including geometric,
diffusional, and hydraulic tortuosity. Here I use the diffusional
tortuosity $\tau= \frac{\epsilon D}{D_{\text{eff}}}$, i.e., the ratio of the 
effective diffusivity $D_{\text{eff}}$ and the intrinsic diffusivity
$D$. I calculate the tortuosity using the procedure detailed in Section \ref{section:tortuosity}

Figure \ref{fig:tau_B} shows the results for the tortuosity as a
function of the applied magnetic flux for the three particle aspect
ratios. In the absence a magnetic field, the three aspect ratios lead to
a similar tortuosity around $\tau\approx 1.5$, slightly larger than
the tortuosity predicted by the Bruggeman relation
\(\tau=\epsilon^{-0.5}=1.41\) for equal volume fractions
\(\epsilon=0.5\) of the phases
\cite{bruggeman1935tortuosity, tjaden_origin_2016}. The observed
tortuosity is consistent with simulations of gyroid structures by Luo et
al., who found diffusive tortuosities in the range from 1.48 to 1.73
\cite{luo_macroscopic_2020}. Figure \ref{fig:tau_B} shows that in an
applied magnetic field, the tortuosity of bijels stabilized by
anisotropic particles (\(\alpha\neq1\)) becomes anisotropic as well.
Oblate particles (\(\alpha=0.5\)) lead to an increase of the tortuosity
\(\tau_\parallel\) in the direction of the magnetic field while the
tortuosity in the direction perpendicular to the magnetic field remains
around \(\tau_\parallel\approx1.5\). Conversely, prolate particles
(\(\alpha=2\)) lead to an increase of the tortuosity in the direction
perpendicular to the magnetic field, while the tortuosity in the
parallel direction slightly decreases with increasing field strength. In
both cases, the changes in tortuosity are consistent with the
anisotropic domain size and the alignment of particles with the magnetic
field. For prolate particles, the longer axis is aligned with the
magnetic moment and induces larger domain size and lower tortuosity in
this direction. The results show that the magnetic field has a
noticeable effect on the tortuosity of the bijel. I observe that the
largest change of tortuosity is measured at intermediate field strength
\(\bar{B}=0.5\). This suggests that the competition of interfacial and
magnetic forces and the resulting alignment of particles and liquid
domains induces morphological changes that are more complex than a
monotonic increase or decrease. I therefore turn to investigate the
microstructural changes within the bijel morphology in more detail.

@@

Beyond the length scale, curvature analysis offers insights into the
topological landscape of the bijel, facilitating further
microstructure optimization for various applications.
\cite{reeves_quantitative_2016} I plot the area averaged mean and
gaussian curvature \(H\Sigma^{-1}\) and \(K\Sigma^{-2}\) for bijels
stabilized with oblate, spherical and prolate ellipsoids at the final timestep in Figure
\ref{fig:curvature-vs-B_ss}.

\begin{figure} 
    \centering 
    \includegraphics[scale = 0.4]{figures/results/paper1/curvature-vs-B_ss.png} 
    \caption{Plot of the area averaged mean $\langle H \rangle \Sigma^{-1}$ and Gaussian 
            $\langle K \rangle \Sigma^{-2}$ curvature in the left and right respectively. Each 
            plot contains the $\langle H \rangle \Sigma^{-1}$ and $\langle K \rangle \Sigma^{-1}$ 
            at the final timestep averaged across 3 runs for particles with $\alpha = 0.5, 1, 2$. 
            The mean curvature remains 0 on average, while the Gaussian 
            curvature shows a reduction as the magnetic field strength is increased for 
            ellipsoidal particles.} 
    \label{fig:curvature-vs-B_ss}
\end{figure}

Error bars are represented by the standard
deviation of the three simulations per parameter set performed. The area
averaged mean curvature shows little dependence on the applied field
strength with values around 0 and within error of one another. A value
near or at 0 is expected as the fluid volume fraction is equal and the
particles are neutrally wetting, implying the interface should not have
a preferred direction of curvature. \cite{jinnai_interfacial_2001} For
ellipsoidal particles, the gaussian curvature becomes more negative as
larger field strengths are applied. I also see that the gaussian
curvature becomes more positive as the interfacial area of the particle
increases. The trend of these results are inverse to the interfacial
area of each particle, \(A_{rm}\). This is expected behavior and is in
agreement with experiments that have investigated the effect of particle
size on the curvature of bijels. \cite{reeves_quantitative_2016}

To explain the reduction in $\langle K\Sigma^{-2} \rangle$ as a function of the applied field strength, two mechanisms are proposed. The first is that 
the particle alignment to the magnetic field during coarsening changes the topology of the surface, which upon jamming of ellipsoidal particles is locked in. 
The other proposed mechanism is that the time evolution of spinodal decomposition remains unaffected, but that the interface conforms to the magnetic field 
driven arrangement of particles upon jamming. While I cannot apply the Gauss Bonnet theorem to characterize the results, I can still relate the coarsening 
behavior of the bijel through its change in genus to the curvature. The time evolution of the genus serves as a topological technique to characterize the 
coarsening of the bijel, which alongside the $\langle K\Sigma^{-2} \rangle$, allows us to identify deviations from the coarsening behavior of bijels 
stabilized with spherical particles, and with magnetic fields. I calculate the genus by extracting the isosurface of the filled order parameter, where 
$\phi = 0$. It is then meshed using the marching cubes algorithm implemented in scikit-image with a cube size of 1. \cite{van2014scikit} From the vertices(v) 
and faces(f) generated from the marching cubes algorithm, I calculate the number of edges(e). The Euler Poincare characteristic, $\chi = n + f - e$. The genus is 
then obtained from the Euler Poincare parameter as $g = 1 - \chi/2$.

\begin{figure} 
    \centering 
    \includegraphics[scale = 0.4]{figures/results/paper1/genus_curvature_vs_coverage.png} 
    \caption{Plots of Gaussian $\langle K \rangle \Sigma^{-2}$ curvature over time averaged across 3 runs. 
    From left to right, I have particles with $\alpha = 0.5, 1, 2$. I observe that the initial stages of 
    evolution are unaffected by the applied magnetic field. However, after about 10000 timsteps, I see a 
    magnetic field dependence on the evolution of the Gaussian curvature. The preservation of this divergence 
    occurs when ellipsoidal particles are used but not spherical particles.} 
    \label{fig:curvature_vs_coverage}
\end{figure}

The time evolution of $g \Sigma^{-1}$ of bijels stabilized with spherical particles is unaffected by the application of a magnetic field. However, a 
slight decrease in the genus as a function of the applied field is observed for bijels stabilized with ellipsoidal particles. This suggests that the 
evolution of the surface corresponding to the interface is changed by the adsorption and orientation of magnetically responsive ellipsoidal particle. 
I see effects of this change in $\langle G \rangle \Sigma^{-2}$ where I see the evolution of the curvature being delayed, indicated as a rightward 
shift of the plots as a function of the applied field.

Despite all our curvature calculations being normalized with their surface area to volume ratio, I still see curvature differences as a function of the applied 
magnetic field. Experiments have identified that curvature changes can occur due to the formation of monogels, which are particle networks formed from bijels that 
can withstand the remixing of their constituent liquids. \cite{sanz_colloidal_2009, lee_making_2013} This process cannot occur in our simulations as I am utilizing 
electrically neutral particles that only interact through contact forces. Therefore, the change in curvature arises from microstructure changes driven by particle 
alignment to the magnetic field direction. These results provide support for the mechanism of control being due to particle alignment to the magnetic field during 
coarsening changes the topology of the surface, which upon jamming of ellipsoidal particles is locked in. As our particles only interact through contact forces, 
the only mechanism present that can change the topology of the interface would be particle interactions upon contact with each other. Particle interactions are 
controlled through the orientation of particles and the local ordering of particles on the interface. Seeing that there are differences in the gaussian curvature of the bijel,
the domain anisotropy characterized in Figure \ref{fig:D2a_B} may be able to be characterized as changes in the topology of the surface. Chan and Thornton introduced a 
topological technique to characterize the connectivity of surface as a function of a distance from the interface to calculate the channel size distribution. 
\cite{chan_channel_2012} I can use the channel size distribution to investigate how the connectivity of the microstructure and thus the 
channel size is affected as a function of the applied magnetic field.

\begin{figure} 
        \centering 
        \includegraphics[scale = 0.4]{figures/results/paper1/CSD_ss.png} 
        \caption{Plots of the channel size distribution for bijels stabilized by oblate, spherical and oblate particles at the final timestep plotted
                 against the distance from the interface normalized with the characteristic length scale. The channel sizes obtained are averaged over 
                 three runs. I see that the peak locations are shifted to larger and smaller positions for bijels stabilized with ellipsoidal particles
                 upon field application} 
        \label{fig:CSD_B_ss}
    \end{figure}

While run to run variation is observed, the combined data of the channel size distribution shown in Figure \ref{fig:CSD_B_ss} shows some differences 
in the channel size distribution in each bijel stabilized with ellipsoidal particles. It was expected that the domain anisotropy obtained using second moment of the 
structure factor would manifest as two peaks in the channel size distribution. \cite{karthikeyan_formation_2024} This behavior is observed in the channel size 
distribution for bijels stabilized with ellipsoidal particles with $\bar{B} = 1$, but is more difficult to see when smaller magnetic fields are applied. 
These differences arise from the particle ordering and local distribution changes observed.

@@

To shed light on the mechanisms involved in the coarsening dynamics,
I now turn to the kinetics of bijel formation in more detail. The
coarsening dynamics can be characterized by a coarsening speed. Here,
I use the finite difference of the directional domain size to
calculate the components of the coarsening velocity
%
\begin{equation}
u_{L_\beta}(t) = \frac{L_\beta(t+\Delta t_s)-L_\beta(t-\Delta t_s)}{2\Delta t_s} ,
\end{equation}
%
where \(\Delta t_s\) is the time between simulation
snapshots, and $\beta$ denotes either the direction parallel or perpendicular to the magnetic field.

@@

Figure \ref{fig:coarsening_velocity} shows the coarsening speed in the directions parallel and perpendicular to the applied magnetic field. 
The data confirms that for spherical particles, domain coarsening remains isotropic. I observe a higher coarsening speed initially which slows 
at later times when the particles begin to jam. This indicates that the
coarsening is subject to different time scales, as reported previously
by Harting and co-workers \cite{gunther_timescales_2014}. Reeves et
al.~\cite{reeves_particle-size_2015} have pointed out that bijel
formation hinges on the jamming time in relation to the disruption time, i.e., the time scale at which domain pinch-off events can cause bijel break-up through 
depercolation. In our simulations, however, I have not observed disruption of bijel formation; stable bijels form for all parameters consistent with simulations 
by Stratford et
al.~\cite{stratford_colloidal_2005} and Jansen et al.~\cite{jansen_bijels_2011}.

For ellipsoidal particles, our data clearly shows that an\-isotropic
domain coarsening arises in magnetic fields.  For oblate particles
(\(\alpha=0.5\)), the coarsening speed in the direction perpendicular
to the field increases with increasing magnetic field strength.
Jamming in this direction appears to be delayed compared with the
direction parallel to the field. The coarsening speed in the parallel
direction remains comparable to the case without an applied magnetic
field. This anisotropic behavior of the coarsening speed reverses for
prolate particles, where the coarsening speed in the direction parallel to the field increases with increasing field strength, and jamming in this direction
is delayed compared with the perpendicular direction. In addition, I observe a shoulder (around \(10^4\Delta t\)) where the coarsening speed decays more slowly before jamming sets in. These results suggest that the coarsening near the jamming time is affected by a mechanism that depends on the applied magnetic field and causes the anisotropic
behavior. This mechanisms involves the re-orientation of anisotropic
particles and their alignment relative to the direction of the
magnetic field. To confirm this hypothesis, I analyze the
orientational order of the particles and their alignment relative to
the magnetic field and the interface, respectively.

@@

I select $\langle Q6 \rangle$ and $\langle Q2 \rangle$ as the bond order parameters to calculate as I am interested in identifying the point when ordering 
of the material changes. Kapfer et al. demonstrated that only using $\langle Q6 \rangle$ to identify the ordering of particles in the system can result in 
false positives when the particles have icosahedral or amorphous structure. \cite{kapfer_jammed_2012} Mickel et al. showed how increasing deviations from 
$\langle Q2 \rangle = 0$ can be used to characterize increasing disorder in a system. \cite{mickel_shortcomings_2013} By plotting $\langle Q6 \rangle$ and 
$\langle Q2 \rangle$, I will be able to identify changes in the local ordering of the system even if I cannot directly assess the local structure quantitatively.

\begin{figure} 
    \centering 
    \includegraphics[width=\columnwidth]{figures/results/paper1/steinhardt_vs_coverage.png} 
    \caption{Time dependence of the mean two and six fold symmetry parameter, $\langle Q2 \rangle$ and $\langle Q6 \rangle$ 
    for bijels stabilized with oblate(left), spherical(middle) and prolate(right) particles. I use $\langle Q2 \rangle$ as an
    indicator of global particle translational order while $\langle Q6 \rangle$ is an indicator of local order.} 
    \label{fig:steinhardt_coverage} 
\end{figure}
